---
title: "The Impact of Recreational Marijuana Legalization on Binge Drinking in the U.S.: Evidence from a Difference-in-Differences Approach"
output: 
  html_document:
    css: serif.css
date: "2025-03-30"
---

## Setup

```{r setup, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
packages <- c("did", "ggplot2", "dplyr", "readr", "stargazer", "showtext")

#install packages if not installed
installed <- packages %in% rownames(installed.packages())
if (any(!installed)) {
  install.packages(packages[!installed])
}

#load packages
lapply(packages, library, character.only = TRUE)

theme_serif <- function(base_size = 14) {
  theme_minimal(base_family = "serif", base_size = base_size) +
    theme(
      text = element_text(family = "serif", size = base_size),
      legend.text = element_text(size = base_size * 0.9),
      axis.title = element_text(size = base_size),
      plot.title = element_text(size = base_size * 1.3, face = "bold"),
      plot.subtitle = element_text(size = base_size)
    )
}


```

## Importing Data

```{r import, echo=TRUE, message=FALSE, warning=FALSE}
url <- 'https://docs.google.com/spreadsheets/d/e/2PACX-1vS-Qiz6Osc9I39bvo_94Mo1GvYe_ghAVboqYvR5xu6ZSPDI-ESb6ECCXgocUP3ALVQ3oxi_n3wDihQL/pub?gid=1277276287&single=true&output=csv'
data <- read.csv(url)
head(data)
summary(data)
```

## Latex Summary Statistics Table
```{r summary}
#subset to key variables only
summary_vars <- data[c("Binge_Drinking_Prevalence", "Legalized", "Bachelors_Rate", "Median_Age", "Urbanization_Rate")]

#create summary table in LaTeX format (default)
stargazer(summary_vars, type = "latex", title = "Summary Statistics", digits = 2, summary.stat = c("min", "mean", "sd", "max"))
```

## Analysis


### Event Study Estimates of Binge Drinking Prevalence by Legalization Cohort
```{r by_year_analysis, echo=TRUE, message=FALSE, warning=FALSE}

att_did <- att_gt(
  yname = "Binge_Drinking_Prevalence",
  tname = "Year",
  idname = "State_ID",
  gname = "G",
  xformla = ~ Bachelors_Rate + Median_Age + Urbanization_Rate,
  data = data,
  est_method = "dr"
)

summary(att_did)
ggdid(att_did) + theme_serif()
```


### Trends in Binge Drinking Prevalence by Legalization Timing
```{r by_year_line_plot, echo=TRUE, message=FALSE, warning=FALSE}

#treated states by year (2012, 2014, 2016)
treated_data <- data %>%
  filter(!is.na(Legalization_Year)) %>%
  mutate(Cohort = as.factor(Legalization_Year))

#control group
control_data <- data %>%
  filter(is.na(Legalization_Year)) %>%
  mutate(Cohort = "Control")

#combine groups
combined_data <- bind_rows(treated_data, control_data)

#average binge drinking by group and year
avg_trends <- combined_data %>%
  group_by(Year, Cohort) %>%
  summarise(Avg_Binge = mean(Binge_Drinking_Prevalence, na.rm = TRUE), .groups = "drop")

#get legalization years per cohort (excluding control)
cohort_lines <- treated_data %>%
  distinct(Cohort, Legalization_Year)

#plot
ggplot(avg_trends, aes(x = Year, y = Avg_Binge, color = Cohort)) +
  geom_line(size = 1.1) +
  geom_point(size = 2) +
  #add vertical line for each treated cohort
  geom_vline(data = cohort_lines, aes(xintercept = Legalization_Year, color = Cohort),
             linetype = "dashed", show.legend = FALSE) +
  labs(
    title = "Binge Drinking Trends by Legalization Cohort and Control Group",
    subtitle = "Dashed lines mark legalization years for each cohort",
    x = "Year",
    y = "Avg. Binge Drinking Prevalence (%)",
    color = "Group"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom") + theme_serif()

```

### Group-Time ATT Estimates of Legalization Effects
```{r grouped, echo=TRUE, message=FALSE, warning=FALSE}
library(broom)
#convert att_gt object into a tidy dataframe
gt_df <- tidy(att_did)

ggplot(gt_df, aes(x = time, y = estimate, color = as.factor(term), group = term)) +
  geom_line(linewidth = 1.1) +
  geom_point(linewidth = 2) +
  geom_errorbar(aes(ymin = estimate - 1.96 * std.error, ymax = estimate + 1.96 * std.error),
                width = 0.3, alpha = 0.6) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  scale_color_brewer(palette = "Dark2", name = "Treatment Year") +
  labs(
    title = "Event Study: Group-Time ATT Estimates",
    subtitle = "Callaway & Sant'Anna (2021)",
    x = "Calendar Year",
    y = "Treatment Effect (ATT)"
  ) +
  theme_minimal(base_size = 14) + theme_serif()
```

### Effect of Length of Exposure to Legalization
``` {r exposure_length, echo=TRUE, message=FALSE, warning=FALSE}
#aggregate plotting code
agg_event <- aggte(att_did, type = "dynamic")
ggdid(agg_event) + theme_serif()
```

### Overall Estimated ATT
```{r overall, echo=TRUE, message=FALSE, warning=FALSE}
#overall ATT

agg_overall <- aggte(att_did, type = "simple")
summary(agg_overall)

overall_df <- data.frame(
  Label = "Overall ATT",
  Estimate = agg_overall$overall.att,
  Lower = agg_overall$overall.att - 1.96 * agg_overall$overall.se,
  Upper = agg_overall$overall.att + 1.96 * agg_overall$overall.se
)

ggplot(overall_df, aes(x = Label, y = Estimate)) +
  geom_col(fill = "steelblue", width = 0.5) +
  geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0.1) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  labs(
    title = "Overall ATT: Effect of Marijuana Legalization",
    subtitle = "Estimated impact on binge drinking prevalence",
    y = "Estimated ATT",
    x = ""
  ) +
  theme_minimal(base_size = 14) + theme_serif()
```

### Treated vs. Control Trends in Binge Drinking Prevalence
```{r overall_trend, echo=TRUE, message=FALSE, warning=FALSE}

#create average trends by treatment status
avg_trends <- data %>%
  group_by(Year, Legalized) %>%
  summarize(
    avg_binge = mean(Binge_Drinking_Prevalence, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Legalized = ifelse(Legalized == 1, "Treated (Legalized)", "Control (Not Legalized)")
  )

#plot
ggplot(avg_trends, aes(x = Year, y = avg_binge, color = Legalized)) +
  geom_line(size = 1.1) +
  geom_point(size = 2) +
  scale_color_manual(
    values = c("Treated (Legalized)" = "darkgreen", "Control (Not Legalized)" = "gray60")
  ) +
  labs(
    title = "Average Binge Drinking Prevalence Over Time",
    subtitle = "Comparing Treated vs. Control States",
    x = "Year",
    y = "Average Binge Drinking Prevalence (%)",
    color = "Group"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom") + theme_serif()
```

### State-Level Trends in Binge Drinking Prevalence by Legalization Status
```{r individual_trends, echo=TRUE, message=FALSE, warning=FALSE}
ggplot(data, aes(x = Year, y = Binge_Drinking_Prevalence, group = State, color = as.factor(Legalized))) +
  geom_line(alpha = 0.5, size = 0.4) +
  scale_color_manual(
    values = c("0" = "darkgray", "1" = "darkgreen"),
    labels = c("0" = "Not Legalized", "1" = "Legalized"),
    name = "Legal Status"
  ) +
  labs(
    title = "Binge Drinking Prevalence Over Time by State",
    subtitle = "Lines represent individual states; color indicates legalization status",
    x = "Year",
    y = "Binge Drinking Prevalence (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom") + theme_serif()
```
